---
title: "First Assignement"
author: "Davide Delfino, Elena Mannino, Sara Tozzi, Marco"
date: "2024-10-03"
output:
  pdf_document:
    toc: yes
    latex_engine: xelatex
  html_notebook:
    number_sections: yes
    toc: yes
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Calibri Light}
subtitle: 
editor_options:
  chunk_output_type: inline
---
Commands to set the directory
```{r}
getwd()
path <- "C:/Users/Davide/Desktop/Alma Mater/SECOND YEAR/GIS/Assignment"
setwd(path)
```
Those are the libraries we will need to fulfill each task
```{r}
library(sf)
library(spData)
library(tidyverse) 
library(ggplot2)
library(readxl)
library(dplyr)
library(viridis)
library(exactextractr)
library(cowplot)
library(ggpubr)
library(patchwork)
library(gridExtra)
```

# BRAZIL *Road Infrastructure*

**Paper**: Morten, M. & Oliveira, J., 2018. *The Effects of Roads on Trade and Migration: Evidence from a Planned Capital City*

**Aim**: Replicate Figure 1 from the paper, which illustrates *Brazil’s capital road infrastructure*.

Import data set on Brasil roads, alongside impose a simple structure to allow for reading the shape file. Since we'll also need cities in our final plot, let's import also an excel file with brazilian cities.
```{r}
br_roads <- "data/SNV_202001A.shp"
sf.br_roads <- st_read(br_roads)
br_cities <- read_excel("data/br.xlsx")
```

After eliminating not available observations
```{r}
br_cities <- br_cities %>%
  filter(!is.na(lng), !is.na(lat))
br_cities_sf <- st_as_sf(br_cities, 
                                 coords = c("lng", "lat"), 
                                 crs = 4326)
br_cities_labels <- br_cities_sf %>%
  st_coordinates() %>%
  as.data.frame() %>%
  mutate(city = br_cities_sf$city)
```

Plot Brazil's road infrastructure
```{r}
ggplot() +
  geom_sf(data = sf.br_roads, aes(color = leg_multim), size = 0.5) +  
  geom_sf(data = br_cities_sf, shape = 21, fill = "black", color = "black", size = 0.5) +
  geom_text(data = br_cities_labels, aes(x = X, y = Y, label = city),
            nudge_y = 0.1, color = "black", size = 3.5, check_overlap = TRUE) +
  labs(title = "Federal Highways and Cities in Brazil",caption = "Brasil") +
  theme_minimal() +
  theme(legend.position = "right", 
        plot.caption = element_text(hjust = 0.5, size = 10),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) 

```

Clean the enviorment for the next Assignment
```{r}
rm(list=ls())
```

# BRAZIL *meso-regions population density*

**Paper**: Pellegrina, H.S. and Sotelo, S., 2021. *Migration, Specialization, and Trade: Evidence from Brazil’s*
*March to the West*

**Aim**: Replicate Figure 2 from the paper, which illustrates *Population in Brazil’s meso-regions (or districts) in different periods*

Import data set on Brasil roads and population. Recall to turn into sf the shape file. 
*Attention*:The original Excel file did have formatting issues, such as empty cells in the first row, which could affected the import process. Due to this issues some adjustment to the original dataset is been performed
```{r}
br_mesoregiones <-"data/BR_Mesorregioes_2021.shp"
sf.br_mesoregiones <- st_read(br_mesoregiones)
pop_mesoregiones <- read_excel("data/BrazilPopulation.xlsx")
```
Before plotting the whole density we need to identify Western regions and aggregate them togheter using the function st_union(). One way is to create manually a new variable to differenciate West and East. To do that we have visualized the location of each region SIGLA in a plot using ggplot. 
```{r}
West <- c("AC", "AM", "AP", "GO", "MA", "MT", "MS", "PA", "TO","RO", "RR")
East <- c("AL", "BA", "CE", "DF", "ES", "MG", "PB", "PE", "PI", "PR", "RJ", "RN", "RS", "SC", "SE", "SP")

sf.br_mesoregiones <- sf.br_mesoregiones %>%
  mutate(regiones = case_when(
    SIGLA %in% West ~ "West",
    SIGLA %in% East ~ "East",
    TRUE ~ "Altro"
  ))

west_regions <- sf.br_mesoregiones[sf.br_mesoregiones$regiones == "West", ]
west_union <- st_union(west_regions) # it is a Multipolygon but we need to turn it in a simple feature
west_union_sf <- st_sf(geometry = st_union(west_union))
```

The main focus in the Paper is on the western meso-regions, as we plot in the following picture
```{r}
ggplot() +
  geom_sf(data = sf.br_mesoregiones, aes(colour = NA)) +
  geom_sf(data = west_union_sf, fill = NA, color = "red") +  
  theme_minimal() +
  labs(color = "Western meso-regiones' population density") 
```
Sinse we do not have data on meso-regiones population density we left_join the original dataset on Brazil with a dataset that still doesn't have any info on density, but at least collects population's data.
```{r}
Brazil <- sf.br_mesoregiones %>% 
  left_join(pop_mesoregiones,by = c("NM_MESO" =  "Brasil, Grande Região, Unidade da Federação e Mesorregião Geográfica"))
```

Let's try to create three plots on the evolution of brazil's population density in 30-years period.
convert in factor years and compute the density for each year. Sinice in every year observations on popiulation is missing forn Lago mirim and Lago dos patos, we can exclude them from the dataset
```{r}
# Rimuovi le osservazioni con NA nella colonna "1991"
Brazil <- Brazil %>%
  filter(!is.na(`1991`))  # Usa backticks se il nome della colonna contiene numeri
```

calcoliamo la "densità di popolazione" del brasile dopo aver convertito in numeriche le variabili di interesse
```{r}
Brazil$`1991` <- as.numeric(as.character(Brazil$`1991`))
Brazil$`2000` <- as.numeric(as.character(Brazil$`2000`))  
Brazil$`2010` <- as.numeric(as.character(Brazil$`2010`))  
```

Total population by year of each meso region over the mean 
```{r}
pop_mean_1991 <- mean(Brazil$`1991`, na.rm = TRUE)
pop_mean_2000 <- mean(Brazil$`2000`, na.rm = TRUE)
pop_mean_2010 <- mean(Brazil$`2010`, na.rm = TRUE)
```

create a new variable with density. Compute for every region the percentage population density
```{r}
Brazil <- Brazil %>%
  mutate(
    density_1991 = `1991` / pop_mean_1991,  
    density_2000 = `2000` / pop_mean_2000,
    density_2010 = `2010` / pop_mean_2010
  )
```
-4 meso-regiones:  hanno una popolazione superiore a 4 volte la media
-5 mesoregiones: hanno una popolazione superiore a 2 volte la media
Creiamo manualmente delle categorie basate sui valori di densità per ogni anno ed assegnamo dei colori corrispondenti per ogni categoria
```{r}
Brazil <- Brazil %>%
  mutate(density_category_1991 = cut(density_1991, 
                                 breaks = c(-Inf, 0.5, 1, 2, 4, Inf), 
                                 labels = c("Very Low", "Low", "Medium", "High", "Very High"),
                                 right = FALSE))


Brazil <- Brazil %>%
  mutate(density_category_2000 = cut(density_2000, 
                                 breaks = c(-Inf, 0.5, 1, 2, 4, Inf), 
                                 labels = c("Very Low", "Low", "Medium", "High", "Very High"),
                                 right = FALSE))


Brazil <- Brazil %>%
  mutate(density_category_2010 = cut(density_2010, 
                                 breaks = c(-Inf, 0.5, 1, 2, 4, Inf), 
                                 labels = c("Very Low", "Low", "Medium", "High", "Very High"),
                                 right = FALSE)) 


colors <- c("Very Low" = "#f0f8ff",  # Azzurro chiarissimo
            "Low" = "#87ceeb",      
            "Medium" = "#4682b4",  
            "High" = "blue",     
            "Very High" = "navy") # Blu scuro
```

creiamo un grafico per ogni anno di interesse( sono diversi da quelli del paper per via di dataset differneti)
```{r}
legend_title <- "pop.share"

# Primo grafico per il 1991
plot_1991 <- ggplot() +
  geom_sf(data = Brazil, aes(fill = density_category_1991)) +
  geom_sf(data = west_union_sf, fill = NA, color = "red", size = 0.5) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(legend.text = element_text(size = 6), 
        legend.title = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = c(0, 0),
        legend.justification = c(0, 0), 
        axis.title = element_blank(),  
        axis.text = element_blank(),    
        axis.ticks = element_blank()) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5,title = legend_title)) + 
  ggtitle("1991") + theme(plot.title = element_text(hjust = 0.5))

# Secondo grafico per il 2000
plot_2000 <- ggplot() +
  geom_sf(data = Brazil, aes(fill = density_category_2000)) +
  geom_sf(data = west_union_sf, fill = NA, color = "red", size = 0.5) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(legend.text = element_text(size = 6), 
        legend.title = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = c(0, 0),
        legend.justification = c(0, 0),
        axis.title = element_blank(),  
        axis.text = element_blank(),    
        axis.ticks = element_blank()) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5,title = legend_title)) + 
  ggtitle("2000") + theme(plot.title = element_text(hjust = 0.5))

# Terzo grafico per il 2010
plot_2010 <- ggplot() +
  geom_sf(data = Brazil, aes(fill = density_category_2010)) +
  geom_sf(data = west_union_sf, fill = NA, color = "red", size = 0.5) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(legend.text = element_text(size = 6), 
        legend.title = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = c(0, 0),
        legend.justification = c(0, 0),
        axis.title = element_blank(),  
        axis.text = element_blank(),    
        axis.ticks = element_blank()) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5,title = legend_title)) + 
  ggtitle("2010") + theme(plot.title = element_text(hjust = 0.5))

# Combinazione dei grafici
combined_plot <- plot_1991 + plot_2000 + plot_2010 + plot_layout(ncol = 3)

# Titolo unico
final_plot <- combined_plot + 
  plot_annotation(title = "The Spatial Distribution of the Brazilian Population between 1950 and 2010", 
                  theme = theme(plot.title = element_text(size = 13, hjust = 0.5)))

print(final_plot)
```

Clean the enviorment for the execution of the next task
```{r}
rm(list=ls())
```

# SOUTH AFRICA

**Paper**: Mettetal, E., 2019. *Irrigation dams, water and infant mortality: Evidence from South Africa*

**Aim**: Replicate Figure 2 from the paper, which illustrates *hydro dams in South Africa*

Load datasets
```{r}
dams.shp <- "data/mygeodata_merged.shp"
sf.dams <- st_read(dams.shp)

SouthAfrica.shp <- "data/LocalMunicipalities2018_Final.shp"
sf.SouthAfrica <- st_read(SouthAfrica.shp)
```


Load river gradient raster
```{r}
#install.packages(c("raster", "sf", "exactextractr"))
library(raster)
```

```{r}
river_gradient_raster <- raster("data/South_Africa_SRMT30meters.tif")
```

Ensure CRS alignment between shapefiles and raster
```{r}
sf.SouthAfrica <- st_transform(sf.SouthAfrica, crs(river_gradient_raster))

```

Add the extracted river gradient data to the shapefile
```{r}
river_gradients_by_region <- exact_extract(river_gradient_raster, sf.SouthAfrica, 'mean')

```

Extract average river gradient for each region
```{r}
sf.SouthAfrica$avg_river_gradient <- river_gradients_by_region
st_crs(sf.SouthAfrica) == crs(river_gradient_raster)
plot(river_gradient_raster)
plot(sf.SouthAfrica$geometry, add=TRUE)
```


Check CRS of the raster (optional)
```{r}
crs(river_gradient_raster)

```

Check CRS of the vector data (optional)
```{r}
st_crs(sf.SouthAfrica)

```

Remove NA
```{r}
sf.SouthAfrica <- sf.SouthAfrica %>% filter(!is.na(avg_river_gradient))

```

Classify the average river gradient into 6 intervals
```{r}
sf.SouthAfrica <- sf.SouthAfrica %>%
  mutate(gradient_class = cut(avg_river_gradient, 
                              breaks = 6, 
                              labels = c("0.0000 - 2.229103", 
                                         "2.229103 - 3.821713", 
                                         "3.821713 - 6.593167", 
                                         "6.593167 - 9.879707", 
                                         "9.879707 - 13.130823", 
                                         "13.130823 - 19.829017"),
                              include.lowest = TRUE))

```

Create the plot
```{r}
ggplot() +
  geom_sf(data = sf.SouthAfrica, aes(fill = gradient_class), color = "white") + 
  geom_sf(data = sf.dams, size = 0.2, color = "black") +
  scale_fill_manual(values = c("lightgrey", "#d9d9d9", "#bdbdbd", "#969696", "#636363", "#4d4d4d"),
                    name = "Avg River Gradient") + 
  theme_minimal() +
  labs(title = "Average River Gradient by Region in South Africa",
       fill = "Avg River Gradient") +
  theme(
    legend.position = c(0, 1),
    legend.justification = c(0, 1), 
    legend.direction = "vertical",
    legend.key.size = unit(0.5, "cm"), # Make the legend keys smaller
    legend.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```
We get a slightly different plot with respect to the figure we want to replicate. 
Differences are due to the datasets. We did not find the same District boundaries
and this probably caused the average river gradient to be computed over different 
regions in terms of dimension.


# ETHIOPIA *distircts*

**Paper**: Fried, S. and Lagakos, D., 2021. *Rural electrification, migration and structural transformation:*
*Evidence from Ethiopia*

**Aim**: Replicate Figure 2 from the paper, which illustrates *districts and electricity grid in Ethiopia*

Dataset we'll need:
```{r}
#data that ontains data for the major roads of Ethiopia
shproads<-"data/Ethiopia_Roads.shp"
sf.roads<-st_read(shproads)

#data thatcontains the spatial data for Electricity grids across the African continent
shppwr<-"data/africa_grid_20170906final.shp"
sf.power<-st_read(shppwr)

#power plant data
plantsafr2<-read_excel("data/RePP_Petersetal.xlsx", sheet= "S3 HPPD", skip=2)


#population density data
shp3<-"data/eth_weredas.shp"
podens_district<-st_read(shp3)
```

Filter the data to focus only on Ethiopian datas
```{r}
sf.pwret<-sf.power %>%
  filter(country=="Ethiopia")
plantseth<-plantsafr2 %>%
  filter(country=='Ethiopia')
```

usare la funzione st_zm() del pacchetto sf in R per rimuovere la dimensione "Z(density)" dalle geometrie.
```{r}
# remove height (or depth) coordinates from a spatial object
sf.roads <- st_zm(sf.roads)
sf.pwret <- st_zm(sf.pwret)
```

do a log transformation of the population density data to allow for a better visualization.
```{r}
podens_district$log_POP_DENS <- log1p(podens_district$POP_DENS)
```

Assign a CRS (Coordinate Reference System) to the population density to insure compatibility between datasets. Therefore, we assigned the CRS corresponding to Ethiopia to ensure compatibility with the other components of the graph. Similarly, it is necessary to transform the power plant dataframe into an sf object and then set the appropriate CRS.

```{r}
podens_district <- st_set_crs(podens_district, 32637) # Replace with the appropriate EPSG code
plantseth_sf <- st_as_sf(plantseth, coords = c("lon", "lat"), crs = 4326)  # Set to WGS 84 (EPSG:4326)
plantseth_sf <- st_transform(plantseth_sf, 32637)  
st_crs(plantseth_sf)
```

Plot the graph.

In the graph, the yellow point represents the Ethiopian power plants, the red lines indicate the high-voltage grid, and the black lines depict the major roads of Ethiopia.

```{r}
ggplot() +
  geom_sf(data = podens_district, aes(fill = log_POP_DENS)) +  # Fill based on log_POP_DENS
  scale_fill_gradientn(colors = c("lightyellow", "lightblue", "blue", "darkblue"), 
                       name = "Log of Population Density") +  
  geom_sf(data = sf.roads, color = "black", size = 20) +  # Adjust the size for roads
  geom_sf(data = sf.pwret, color = "red", size = 15) +  # Power plants in red
  geom_sf(data = plantseth_sf, color = 'yellow', size = 2) +   # Set point size for visibility
  labs(
    title = "Ethiopian Population Density and The Electric Grid", 
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 10))
```
Clean the enviorment for the execution of the next task
```{r}
rm(list=ls())
```

# VIETNAM *Roads infrastruture*

**Paper**: Balboni, C.A., 2019. In harm’s way? *infrastructure investments and the persistence of coastal cities*

**Aim**: Replicate Figure 2 from the paper, which illustrates *Vietnam’s road infrastructure by road type*

Import data set on Vietam's roads
```{r}
Vietnam_roads <-"data/vnm_rdsl_2015_OSM.shp"
sf.Vietnam_roads <- st_read(Vietnam_roads)
```

Aggregate road type in 5 cathegories
```{r}
sf.Vietnam_roads <- sf.Vietnam_roads %>%
  mutate(road_category = case_when(
    type %in% c("motorway", "motorway_link") ~ "Freeway",
    type %in% c("trunk", "trunk_link") ~ "Dual carriageway",
    type %in% c("primary", "primary_link") ~ "Major roads",
    type %in% c("secondary", "secondary_link", "tertiary", "tertiary_link") ~ "Minor roads",
    TRUE ~ "Other roads"
  ))
```

For some reason the aes(size) doesn't work
```{r}
ggplot() +
  geom_sf(data = sf.Vietnam_roads, aes(size = road_category, color = road_category)) +  
  scale_size_manual(values = c("Freeway" = 3, 
                               "Dual carriageway" = 2.5, 
                               "Major roads" = 2, 
                               "Minor roads" = 1.5, 
                               "Other roads" = 1)) +  
  scale_color_manual(values = c("Freeway" = "red", 
                                "Dual carriageway" = "blue", 
                                "Major roads" = "green", 
                                "Minor roads" = "orange", 
                                "Other roads" = "purple")) +  
  labs(title = "Tipologie di Strade in Vietnam e Città", 
       size = "Spessore delle Strade", 
       color = "Categoria di Strada") +
  theme_minimal()

```

```{r}
ggplot() +
  geom_sf(data = sf.Vietnam_roads, aes(color = road_category)) +  
  scale_color_manual(values = c("Freeway" = "red", 
                                "Dual carriageway" = "blue", 
                                "Major roads" = "green", 
                                "Minor roads" = "orange", 
                                "Other roads" = "purple")) +  
  labs(title = "Vietnam's roads type", 
       color = "Type") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),         # Rimuove la griglia
    legend.position = "left",             # Sposta la leggenda sulla sinistra
    axis.title = element_blank(),          # Rimuove i titoli degli assi
    axis.text = element_blank(),           # Rimuove i testi degli assi
    axis.ticks = element_blank()           # Rimuove i tick marks degli assi
  )

```
ultima prova ma la size non funziona 
```{r}
sf.Vietnam_roads$road_category <- as.factor(sf.Vietnam_roads$road_category)
aes(color = road_category, size = ifelse(road_category == "Freeway", 2, 1))


ggplot() +
  geom_sf(data = sf.Vietnam_roads, aes(color = road_category, size = road_category)) +  
  scale_color_manual(values = c("Freeway" = "red", 
                                 "Dual carriageway" = "blue", 
                                 "Major roads" = "green", 
                                 "Minor roads" = "orange", 
                                 "Other roads" = "purple")) +  
  scale_size_manual(values = c("Freeway" = 10,  # Spessore per le autostrade
                                "Dual carriageway" = 7,  # Spessore per le carreggiate
                                "Major roads" = 4,  # Spessore per le strade principali
                                "Minor roads" = 2,  # Spessore per le strade secondarie
                                "Other roads" = 0.1)) +  # Spessore per le altre strade
  labs(title = "Vietnam's roads type", 
       color = "Type", 
       size = "Relevance") +  # Aggiungi una legenda per la dimensione
  theme_minimal() +
  theme(
    panel.grid = element_blank(),         # Rimuove la griglia
    legend.position = "left",             # Sposta la leggenda sulla sinistra
    axis.title = element_blank(),          # Rimuove i titoli degli assi
    axis.text = element_blank(),           # Rimuove i testi degli assi
    axis.ticks = element_blank()           # Rimuove i tick marks degli assi
  )

```
clean the enviorment
```{r}
rm(list=ls())
```


# OVERLAPPING *overlap between Mexico and Europe*

**Aim**:  create a plot that overlays the Mexican shapefile onto the European shapefile (excluding Russia and Overseas France if necessary)

import "world" dataset 
```{r}
world<- world
```

The data set propose a division for Frence: 'France' e 'French Southern and Antarctic Lands'. Acceding to the geom of Frence we aknowledge that French Polynesia is considered in this observation.
```{r}
sf.France <- world %>%
  filter(name_long %in% 'France')
```

To exclude this overses land we implemented the following procedure: Isolate Frence, extract the list in geometry involving polygons of Polynesia and eliminate it. Could be reasable to move it toward French Southern and Antarctic Lands.
```{r}
France <- world[world$name_long == "France",]

geom_France <- st_geometry(France)[[1]]
geom_France <- geom_France[-1]

st_geometry(France) <- st_sfc(st_multipolygon(geom_France), crs = st_crs(world))
world[world$name_long == "France", ] <- France  # 5. Sostituisci nel dataset originale se necessario

```

We are now left overlapping Mexico and Europe. The first step is to select just observations we are interested to excluding Russian Federation and Polynesia from Europe.
```{r}
sf.EU <- world %>%
  filter(continent == 'Europe' & name_long != 'Russian Federation')

sf.Mexico <- world %>%
  filter(name_long == 'Mexico')
```

Graph with no manual manipulations. We
```{r}
# Creazione del grafico
ggplot() +
  geom_sf(data = sf.EU, aes(fill = continent), color = "black") +
  geom_sf(data = sf.Mexico, aes(fill = "Mexico"), color = "black", inherit.aes = FALSE) +
  scale_fill_manual(values = c("Europe" = "blue"  , "Mexico" = "red")) + 
  theme_minimal() +
  labs(title = "Europe and Mexico")
```
other graphs
```{r}
ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.2) +
  geom_sf(data = sf.Mexico, aes(fill = region_un), color = "black", alpha = 0.2) +
  theme_minimal() +
  labs(title = "Europe and Mexico", fill = "Region") +
  theme(legend.position = "right")  # Nasconde la legenda se non necessaria

```
Verifica del CRS per sf.EU e sf.Mexico Se il CRS è mancante, assegnalo manualmente.
Applica la traslazione al Messico per sovrapporlo all'Europa. Funzione per spostare manualmente le geometrie (traslazione). Applichiamo manualemnte uno spostamento sia verso est +120 che a nord +20
```{r}
st_crs(sf.EU)
st_crs(sf.Mexico)

if (is.na(st_crs(sf.EU))) {
  st_crs(sf.EU) <- 4326
}

if (is.na(st_crs(sf.Mexico))) {
  st_crs(sf.Mexico) <- 4326
}

sf.EU <- st_transform(sf.EU, crs = 4326)
sf.Mexico <- st_transform(sf.Mexico, crs = 4326)
shift_geom <- function(geom, x_shift, y_shift) {
  st_geometry(geom) <- st_geometry(geom) + c(x_shift, y_shift)
  return(geom)
}
sf.Mexico_shifted <- shift_geom(sf.Mexico, x_shift = 120, y_shift = 20)

```

Mexico will now be shifted. Hence, we need to append to this new geometry the correct coordinate reference system compatible with Europe.
```{r}
st_crs(sf.Mexico_shifted)
if (is.na(st_crs(sf.Mexico_shifted))) {
  st_crs(sf.Mexico_shifted) <- 4326  # Assegna WGS 84 a sf.Mexico
}
st_crs(sf.Mexico_shifted)
```

Now overlapping the 2 figures is possible
```{r}
sf.Mexico_shifted
ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Europe with Mexico on Top", fill = "Region") +
  theme(legend.position = "none")

```

Playing with coordinate refrence systems. comparison among different projections
```{r}
st_crs(sf.Mexico_shifted)<- 3857
st_crs(sf.EU)<- 3857
ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +  # Europa con trasparenza
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +  # Messico sopra con trasparenza
  theme_minimal() +
  labs(title = "Europe with Mexico on Top", fill = "Region") +
  theme(legend.position = "none")  # Nasconde la legenda se non necessaria

```
SOVRAPPOSIZIONE
```{r}
st_crs(sf.Mexico_shifted)<- 3035
st_crs(sf.EU)<- 3035
ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +  # Europa con trasparenza
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +  # Messico sopra con trasparenza
  theme_minimal() +
  labs(title = "Europe with Mexico on Top", fill = "Region") +
  theme(legend.position = "none")  # Nasconde la legenda se non necessaria

```
SOVRAPPOSIZIONE
```{r}
st_crs(sf.Mexico_shifted)<- 4326
st_crs(sf.EU)<- 4326
ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +  # Europa con trasparenza
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +  # Messico sopra con trasparenza
  theme_minimal() +
  labs(title = "Europe with Mexico on Top", fill = "Region") +
  theme(legend.position = "none")  # Nasconde la legenda se non necessaria

```
5. Spostamento senza trasformazione
Se i tuoi dati spaziali non sono stati trasformati correttamente in base al CRS, potresti continuare a vedere le stesse geometrie sulla mappa, ma questo non significa che siano state correttamente allineate al sistema di coordinate desiderato.

3. Rappresentazione su scala globale
Se visualizzi dati a livello globale, alcune proiezioni possono distorcere le aree, ma l'effetto potrebbe non essere evidente se le geometrie sono simili. Le proiezioni come Mercator possono distorcere le dimensioni delle terre vicine ai poli, ma non in modo tale da influenzare drasticamente aree più piccole.

FINAL COMPARISON SCEGLIETE VOI QUALI VOLETE TENERE E PROVATE ALTRE PROJECTIONS
```{r}

st_crs(sf.Mexico_shifted) <- 4326
st_crs(sf.EU) <- 4326
plot_4326 <- ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +
  theme_minimal() +
  labs(subtitle = "CRS: 4326") +
  theme(legend.position = "none") +  # Nasconde la legenda
  coord_sf(expand = FALSE)  # Elimina lo spazio extra

st_crs(sf.Mexico_shifted) <- 3857
st_crs(sf.EU) <- 3857
plot_3857 <- ggplot() +
  geom_sf(data = sf.EU, aes(fill = region_un), color = "black", alpha = 0.5) +
  geom_sf(data = sf.Mexico_shifted, aes(fill = region_un), color = "black", alpha = 0.5) +
  theme_minimal() +
  labs(subtitle = "CRS: 3857") +
  theme(legend.position = "none") +  # Nasconde la legenda
  coord_sf(expand = FALSE)  # Elimina lo spazio extra

# Affiancamento dei due grafici
combined_plot <- grid.arrange(plot_4326, plot_3857, ncol = 2)

# Aggiunta del titolo generale
final_plot <- annotate_figure(combined_plot,
                              top = text_grob("Comparison between different projections", face = "bold", size = 14))

# Visualizzazione del grafico finale
print(final_plot)

```

