---
title: "NetCDF_NASA"
author: "Marco Caravella, Davide Delfino, Elena Mannino"
date: "2024-10-11"
output: html_document
---
Commands to set the directory:
```{r}
getwd()
path <- "C:/Users/Davide/Desktop/Alma Mater/SECOND YEAR/GIS/Assignment"
setwd(path)
```

Those are the libraries we will need to fulfill each task:
```{r}
library(sf)
library(spData)
library(tidyverse) 
library(ggplot2)
library(readxl)
library(dplyr)
library(viridis)
```

# NetCDF_NASA assignment

*Paper:* Porteous, O., 2019. High trade costs and their consequences: An estimated dynamic model of African
agricultural storage and trade. American Economic Journal: Applied Economics, 11(4), pp.327-66.

*Tasks:* download the paper’data, combine with world and transportation data to:
1- Generate an sf MULTIPOINT feature of markets across Africa and plot it;
2- Calculate minimum distance of each market to the (i) nearest road, and (ii) nearest port, (iii) nearest
airport;
3- Produce 3 scatter plots relating market average prices to the minimum distances.

*Note*: the paper appendix contains a lot of useful information about the market price data

# DISCUSSION ON THE PAPER


# DATA PREPARING PROCESS


# MARKET ACROSS AFRICA


# COMPUTATION OF DISTANCES


# RELATION BETWEEN MARKET AVERAGE PRICE AND MINIMUM DISTANCES
# 