---
title: "NetCDF_NASA"
author: "Davide Delfino, student ID 1126595"
date: "2024-11-09"
output:
  pdf_document:
    toc: yes
    latex_engine: xelatex
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Calibri Light}
subtitle: NASA Earthdata Webinar
editor_options:
  chunk_output_type: inline
---

Commands to set the directory:
```{r}
getwd()
path <- "C:/Users/david/Desktop/University/GIS/Exam_SRB"
setwd(path)
```

Those are the libraries we will need to fulfill each task:
```{r}
# Data Wrangling and Transformation
library(tidyverse) 
library(readxl)
library(dplyr)

# Spatial Data Handling and Manipulation
library(sf) # for Vector Data
library(nngeo)
library(spData)
library(terra) # for Raster Data
library(ncdf4)
library(spdep)
library(exactextractr)

# Data Visualization
library(ggplot2)
library(viridis)
library(ggimage)

# conflicts in terra
library(conflicted)
conflict_prefer("extract", "terra")
conflict_prefer("filter", "dplyr") 
```

# Final Exam Structure

Section A: Research Proposal (450 words max)
           - Research Question
           - Data & Methods
           - Expected Outcomes/Contribution
           
Section B: Task replication on Serbian data
           - Mandatory tasks (19 tasks)
           - First not mandatory task
           - Second not mandatory task
           - Third not mandatory task

# SECTION A : RESEARCH PROPOSAL

The research proposal is available in a related PDF. 
The title of the analysis is: "A Monetary Policy Model for forecasting Optimal Euro Area Expansion"

In the following lines I will introduce the research idea using the abstract of the PDF.

The research proposal outlines the development of a monetary policy model to analyze
the responses of European countries outside the Economic and Monetary Union (EMU) to
potential accession. The primary objective is to predict which countries could be optimally
integrated, considering their responses to the shock —namely, the accession itself— and the
endogenous mechanisms suggested by Optimum Currency Area (OCA) theory toward longrun dynamic equilibria. The model will incorporate structural VAR and machine learning
approaches, alongside Geographic Information Systems (GIS) for enhanced visualization and
spatial econometric analysis.


           
# SECTION B : MANDATORY TASKS

## Task 1

Load the shapefile of countries and keep just Serbia
```{r}
shp <- read_sf("data/Countries/ne_10m_admin_0_countries.shp") 
shp <- shp %>%
  filter( SU_A3 == "SRB")
```

Load the shapefile of European NUTS 3 regions and keep just the NUTS 3 of Serbia
```{r}
shp3 <- read_sf("data/NUTS3/NUTS_RG_01M_2021_4326_LEVL_3_repaired.shp")
shp3 <- shp3 %>%
  filter( CNTR_CODE == "RS") # ISO 3166-1 alpha-2 code for Serbia
```

Create a boundary box for Serbia with a small margin
```{r}
box <- st_bbox(shp)
box <- box+c(-.5,-.5,.5,.5)

box_sf <- st_as_sfc(box)
```


Plotting Serbia and the division according to the NUTS 3
```{r}
ggplot() +
  geom_sf(data = shp, fill = "lightblue", color = "grey40", size = 0.2) +
  geom_sf(data = shp3, fill = "palegreen", color = "grey30", size = 0.2) + 
  geom_sf(data = box_sf, fill = NA, color = "red", linetype = "dashed",
          size = 0.8) + 
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    panel.grid = element_blank(), # Rimuove la griglia
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16), 
    plot.caption = element_text(hjust = 0.5, size = 10, face = "italic") 
  ) +
  labs(
    title = " NUTS 3 Serbia",
    caption = "Reference: Natural Earth"
  )  
```
## Task 2

Load SPEI data and crop it around the box
```{r}
spei_data <- rast("data/spei01.nc")
cropped_spei <- crop(spei_data, box)
```

Keep the last six layers assuming that those are the  SPEI values from 2015 to 2020
```{r}
# Check how many layers are in the SPEI data
n_layers <- nlyr(cropped_spei)

# Extract the last six layers (assuming they correspond to 2015-2020)
reduced_spei <- cropped_spei[[ (n_layers-5):n_layers ]]

# Rename these layers sequetinally from 2015 to 2020
names(reduced_spei)
names(reduced_spei) <- c("SPEI_2015", "SPEI_2016", "SPEI_2017", "SPEI_2018",
                         "SPEI_2019", "SPEI_2020")
names(reduced_spei)
```
Plot the spei values from 2015 to 2020
```{r}
par(mfrow = c(2, 3))

for (layer_index in 1:nlyr(reduced_spei)) {
  plot(reduced_spei[[layer_index]], main = names(reduced_spei)[layer_index])
  plot(st_geometry(shp3), add = TRUE, border = "black")
}

par(mfrow = c(1, 1))
```

## Task 3

Fix a seed
```{r}
set.seed(595)
```


Generate a vector of 6 numbers from a uniform distribution (min = 0, max = 1) and multiply the last 6 SPEI layers by these random numbers 
```{r}
random_vector <- runif(6, min = 0, max = 1)

for (i in 1:6) {
  reduced_spei[[i]] <- reduced_spei[[i]] * random_vector[i]
}
```

## Task 4

Extract the grid points of the SPEI raster data
```{r}
points <- as.points(reduced_spei)
points
points_sf <- st_as_sf(points)
```

Plot these points. Note that far the see there is no corresponding point in the grid
```{r}
ggplot()+
  geom_sf(data = shp3) + 
  geom_sf(data = points_sf)
```

## Task 5

Extract the SPEI data for each polygon in “shp” and “shp3”. Keep all considered years.

Methods A and B are both admissible:

A. KEEP ALL YEARS IN A ROW AND VISUALIZE SEPARATELLY THE MEAN VALUES FOR 2015 AND 2020
```{r}
# Extract the mean SPEI data for Serbia, for every year
extracted_spei_shp <- extract(reduced_spei, shp, fun=mean, na.rm=TRUE)
# Extract the mean SPEI data for Serbia's regions, for every year
extracted_spei_shp3 <- extract(reduced_spei, shp3, fun=mean, na.rm=TRUE) 
```

B. KEEP JUST YEARS 2015 AND 2020
```{r}
spei_2015 <- reduced_spei[["SPEI_2015"]]
spei_2020 <- reduced_spei[["SPEI_2020"]]
```

Extraxt spei data from "shp": just Serbia in 2015 and 2020
```{r}
extracted_spei_shp_2015 <- extract(spei_2015, shp, fun = mean, na.rm = TRUE)
extracted_spei_shp_2020 <- extract(spei_2020, shp, fun = mean, na.rm = TRUE)
```

Extraxt spei data from "shp3": all regiones in Serbia in 2015 and 2020
```{r}
extracted_spei_shp3_2015 <- extract(spei_2015, shp3, fun = mean, na.rm = TRUE)
extracted_spei_shp3_2020 <- extract(spei_2020, shp3, fun = mean, na.rm = TRUE)
```

Check the average SPEI level in Serbia in 2015 and in 2020
```{r}
mean_spei_serbia_2015 <- mean(c(extracted_spei_shp_2015$SPEI_2015),
                              na.rm = TRUE)
mean_spei_serbia_2020 <- mean(c(extracted_spei_shp_2020$SPEI_2020),
                              na.rm = TRUE)
mean_spei_NUTSserbia_2015 <- mean(c(extracted_spei_shp3_2015$SPEI_2015),
                                  na.rm = TRUE)
mean_spei_NUTSserbia_2020 <- mean(c(extracted_spei_shp3_2020$SPEI_2020),
                                  na.rm = TRUE)
```

Visualize results
```{r}
cat("Average SPEI in Serbia in 2015:", mean_spei_serbia_2015, "\n")
cat("Average SPEI in Serbia in 2020:", mean_spei_serbia_2020, "\n")
cat("Average SPEI in Serbia in 2015:", mean_spei_NUTSserbia_2015, "\n")
cat("Average SPEI in Serbia in 2020:", mean_spei_NUTSserbia_2020, "\n")
```

COMMENT:  
The Standardized Precipitation-Evapotranspiration Index (SPEI) values calculated represent the average drought/wetness index for Serbia. Positive SPEI values indicate wet conditions, meaning precipitation levels exceed evapotranspiration (higher humidity than average). Although the 202 value is still low, it is slightly more positive than in 2015. This may suggest a slight increase in average humidity in 2020 compared to 2015, though the rise is not significant enough to indicate a true wet climate anomaly.


## Task 6

Load Aqueduct 4.0 dataset 
```{r}
sf_Acqueduct <- read_sf("data/Aqueduct/Aqueduct_baseline.shp")
```

Assume “bws_raw” is the water stress index of each watershed in 2015.

## Task 7

Generate a water stress time series

Set the seed corresponding to the last three values of my student ID
```{r}
set.seed(565) 
```

As said above, assume that “bws_raw” is the water stress index of each watershed in 2015 and add 
the vector to the sf_Aqueduct
```{r}
water_stress_2015 <- sf_Acqueduct$bws_raw 
sf_Acqueduct$water_stress_2015 <- water_stress_2015

# check the lenght of the vector
length(sf_Acqueduct$bws_raw)
```

Generate a matrix of random multipliers for 5 years (2016 to 2020) for 382 elements
```{r}
multipliers <- matrix(rnorm(5 * 382, mean = 1, sd = 1), nrow = 382, ncol = 5)
```

Calculate the annual values by multiplying the 2015 values by the random multipliers
```{r}
water_stress_ts <- water_stress_2015 * multipliers
```

Add those attributes to the Aqueduct shapefile
```{r}
sf_Acqueduct$water_stress_2016 <- water_stress_ts[, 1]
sf_Acqueduct$water_stress_2017 <- water_stress_ts[, 2]
sf_Acqueduct$water_stress_2018 <- water_stress_ts[, 3]
sf_Acqueduct$water_stress_2019 <- water_stress_ts[, 4]
sf_Acqueduct$water_stress_2020 <- water_stress_ts[, 5]
```

## Task 8

Rasterize the values of “water_stress_20XX” and thenextract them to the shapefiles of Serbia and Serbian NUTS 3. 

Define the extent and parameters as the box_sf extension
```{r}
bbox_extent <- ext(vect(box_sf))
years <- 2015:2020  # Years for which to create the rasters
```

Initialize a list to store the rasters
```{r}
rasters <- list()
```

Loop to create and rasterize for each year
```{r}
for (year in years) {
  raster <- rast(bbox_extent, ncol = 1000, nrow = 1000)
  field_name <- paste0("water_stress_", year)
  raster <- rasterize(vect(sf_Acqueduct), raster, field = field_name)
  rasters[[as.character(year)]] <- raster
}
```

Plot each raster with titles for each year
```{r}
par(mfrow = c(2, 3))
for (year in 2015:2020) {
  plot(rasters[[as.character(year)]], main = paste("Water Stress in Serbia -",
                                                   year))
}
par(mfrow = c(1, 1))
```

Extract raster values to Serbia shapefile
```{r}
water_stress_2020_serbia <- exact_extract(rasters[[6]], shp, 'mean')
```

Extract raster values to Serbian NUTS3 shapefile
```{r}
water_stress_2020_nuts3 <- exact_extract(rasters[[6]], shp3, 'mean')
```
What is the average water stress level in Serbia in 2020? Is this value equal to the simple
average of water stress in Serbian NUTS 3 districts? Discuss this shortly.

Calculate average water stress in Serbia for 2020
```{r}
avg_water_stress_2020_serbia <- mean(water_stress_2020_serbia, na.rm = TRUE)
avg_water_stress_2020_serbia
```

Calculate average water stress in Serbian NUTS3 regions for 2020
```{r}
avg_water_stress_2020_nuts3 <- mean(water_stress_2020_nuts3, na.rm = TRUE)
avg_water_stress_2020_nuts3
```


COMMENT:  
I would have expected exactly the same number, but actually they differ 
for a small amount. I believe that the reason for this could arise from some
boundary mismatches between shp and shp3. Serbia’s national boundary may not 
exactly align with the aggregate boundary formed by the NUTS 3 districts.
Moreover Water stress may not be evenly distributed across all NUTS 3 districts, 
causing differences in values when aggregated as a whole versus as smaller districts.


Extract raster values to Serbian NUTS3 shapefile and stor it into columns
```{r}
water_stress_2015_nuts3 <- exact_extract(rasters[[1]], shp3, 'mean')
water_stress_2016_nuts3 <- exact_extract(rasters[[2]], shp3, 'mean')
water_stress_2017_nuts3 <- exact_extract(rasters[[3]], shp3, 'mean')
water_stress_2018_nuts3 <- exact_extract(rasters[[4]], shp3, 'mean')
water_stress_2019_nuts3 <- exact_extract(rasters[[5]], shp3, 'mean')
```

Initialize an empty list to store water stress values for each year
```{r}
years <- 2015:2020
water_stress_list <- list()
```

Loop through each year to calculate mean water stress and store in the list
```{r}
for (i in seq_along(years)) {
  water_stress_list[[i]] <- exact_extract(rasters[[i]], shp3, 'mean')
}
```

Combine the list into a matrix or data frame, with each column named by the respective year
```{r}
water_stress_matrix <- do.call(cbind, water_stress_list)
colnames(water_stress_matrix) <- paste0("water_stress_", years)
```

## Task 9

Load the provided population density data for 2015 and crop it around the boundary 
box already defined. 

Load the data on Serbian population density
```{r}
population_density <- rast("data/gpw-population-density-2015/gpw_v4_population_density_rev11_2015_15_min.asc")
```

Crop them on the box
```{r}
cropped_population_density <- crop(population_density, box)
```


Population density plot
```{r}
plot(cropped_population_density) # high density in Belgrado
plot(shp3$geometry,add = T)
```

What is the average population density level in Serbia? 

compute the average population densiity
```{r}
avg_pop_den <- exact_extract(
  x = cropped_population_density,
  y = shp,
  fun = 'mean')
avg_pop_den # 94.07967 persone per km^2
```


Control the CRS of the two objects before proceding
```{r}
spei_data
cropped_population_density
if (crs(cropped_population_density) != crs(reduced_spei)) {
  cropped_population_density <- project(cropped_population_density,
                                        crs(reduced_spei))
}
```

Resample it on the resolution of SPEI data using the “bilinear” method. 

Risampling
```{r}
resampled_population_density <- resample(cropped_population_density,
                                         reduced_spei, method = "bilinear")
```

Resampled population density plot
```{r}
summary(resampled_population_density)
plot(resampled_population_density)
plot(shp3$geometry,add = T)
```

compute the average population density  after the resampling
```{r}
avg_pop_den_2 <- exact_extract(
  x = resampled_population_density,
  y = shp,
  fun = 'mean')
avg_pop_den_2 # 95.35654 
```

Is the Serbian average population density changed? Discuss this shortly.

COMMENT: 
Yes, the averege population density is changed after the resampling.
This phenomenon could be related to the fact that less the resolution the widther the high populated density area.

# SECTION B : FIRST NOT MANDATORY TASK
create a population-weighted version of SPEI and water stress data using the 
gridded data of population density. 

Calculate the average population density in Serbia
```{r}
pop_density_serbia <- exact_extract(resampled_population_density, shp, 'mean')
avg_pop_density_serbia <- mean(pop_density_serbia, na.rm = TRUE)
```

Create the population density ratio for each grid point
```{r}
pop_density_ratio <- resampled_population_density / avg_pop_density_serbia
```

Initialize a list to store population-weighted SPEI values and plot them all together
```{r}
years <- 2015:2020
pop_weighted_spei <- list()
```

Loop through each year to calculate and plot the population-weighted SPEI
```{r}
par(mfrow = c(2, 3))
for (i in seq_along(years)) {
  pop_weighted_spei[[i]] <- reduced_spei[[i]] * pop_density_ratio
  plot(pop_weighted_spei[[i]], main = paste("Population-weighted SPEI for", 
                                            years[i]))
}
par(mfrow = c(1, 1))
```


## Task 10

Is it relevant to use population-weighted climate variables when analyzing GDP data? Are they relevant when investigating Agricultural GVA or Agricultural yields? And what about Industrial GVA? Discuss this shortly

Using population-weighted climate variables can be relevant in analyzing GDP data, particularly for understanding climate's impact across different sectors. By adjusting climate data based on population distribution, we can better assess how climate affects economic activity, especially in densely populated regions.

In the agricultural sector, where production is often concentrated in specific areas, population-weighting can provide valuable insights into how climatic variations influence productivity and economic outcomes. For instance, understanding how temperature and precipitation affect crop yields in regions with higher population density can inform agricultural policies and economic forecasts.

In contrast, the relevance of population-weighted climate variables in industrial GVA may be less pronounced. Industrial activities often occur in diverse geographical areas that do not necessarily align with population densities, making climate impacts more complex. Industries can be located in regions with varying climatic conditions, thus diminishing the utility of population-weighted analysis.

Overall, while population-weighted climate variables are crucial for understanding agricultural impacts, their relevance for industrial analysis is limited due to the nature of industrial location and climate interaction.

## Task 11

Assume SPEI resolution is c(1,1) and population density resolution is c(0.25,0.25). In your opinion is it preferable to reduce the resolution of SPEI to the resolution of population density and then create a pop-weighted SPEI index, or viceversa? Discuss this shortly.


While higher resolution in population density offers more detail, it risks overfitting when misaligned with climate data, as it can lead to spurious correlations that do not accurately reflect underlying patterns. Reducing the SPEI resolution to match that of population density is preferred because it averages values over relevant spatial units, yielding a more accurate population-weighted SPEI index. This approach ensures that the index better reflects the actual climate impacts on populated areas, allowing for a more meaningful analysis of how climate variability affects population-dense regions. Overall, aligning the resolutions enhances the reliability of the findings and supports more informed decision-making.

## Task 12

Load the Sector A Serbian data (NACE classification)
```{r}
SRB_secA <- read_xlsx("data/Serbian_gva_sector_a.xlsx", sheet = 2)
SRB_secA <- SRB_secA %>%
  filter(TERRITORY_ID != "RSZZZ")%>%
  select( c(TERRITORY_ID,`2015`,`2016`,`2017`,`2018`,`2019`,`2020`))%>%
  rename(
    agr_gva_2015 = `2015`,
    agr_gva_2016 = `2016`,
    agr_gva_2017 = `2017`,
    agr_gva_2018 = `2018`,
    agr_gva_2019 = `2019`,
    agr_gva_2020 = `2020` 
  )
```


Merge them with the sf object called “shp3”
```{r}
shp3 <- shp3 %>%
  left_join(SRB_secA, by = join_by(NUTS_ID == TERRITORY_ID))
```

## Task 13

Generate a new simple feature object called “final_sf” whose features are the 
Serbian NUTS 3 polygons and whose fields are the average values of SPEI,
water stress, and Agricultural GVA in each considered year. 
```{r}
final_sf <- shp3 %>%
  select(NUTS_ID, agr_gva_2015,agr_gva_2016,agr_gva_2017,agr_gva_2018,
         agr_gva_2019, agr_gva_2020,geometry) %>%
  cbind(extracted_spei_shp3, water_stress_matrix) %>%
  select(!ID)
```

compute the avereges for each variable
```{r}
final_sf_avg <- final_sf %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
```

## Task 14 

Plot the time series of SPEI and water stress in Serbia from 2015 to 2020.

Time series variable for SPEI values from 2015 to 2020
```{r}
#Reshape in long form the dataset
final_SPEI_sf_long_avg <- final_sf_avg %>%
  pivot_longer(cols = starts_with("SPEI_"), 
               names_to = "year",
               values_to = "spei") 

# Remove "SPEI_" from the year name to leave just the year
final_SPEI_sf_long_avg <- final_SPEI_sf_long_avg %>%
  mutate(year = as.numeric(gsub("SPEI_", "", year)))
```

Time series plot for SPEI from 2015 to 2020
```{r}
SPEI_avg_time_series <- ggplot(final_SPEI_sf_long_avg,
                               aes(x = year, y = spei)) +
  geom_line(color = "blue") + 
  geom_point(color = "red") +   
  labs(title = "Time Series of SPEI (2015-2020)",
       x = "Year",
       y = "SPEI") +
  theme_minimal()
SPEI_avg_time_series
```

Time series variable for water stress values from 2015 to 2020
```{r}
final_water_sf_long_avg <- final_sf_avg %>%
  select(water_stress_2015, water_stress_2016, water_stress_2017,
         water_stress_2018, water_stress_2019, water_stress_2020) %>%
  pivot_longer(cols = starts_with("water_stress"),
               names_to = "year",   
               values_to = "water_stress")

# Rimuovi "SPEI_" dal nome dell'anno per ottenere solo il numero
final_water_sf_long_avg <- final_water_sf_long_avg %>%
  mutate(year = as.numeric(gsub("water_stress_", "", year)))
```

Time series plot for water stress values from 2015 to 2020
```{r}
water_avg_time_series <- ggplot(final_water_sf_long_avg,
                                aes(x = year,y = water_stress)) +
  geom_line(color = "blue") +     
  geom_point(color = "red") +     
  labs(title = "Time Series of water stress (2015-2020)",
       x = "Year",
       y = "water stress") +
  theme_minimal()
water_avg_time_series
```

Merge the two datasets and generate a combined plot
```{r}
final_combined_long_avg <- final_SPEI_sf_long_avg %>%
  select(year, spei) %>%
  rename(value = spei) %>%
  mutate(variable = "SPEI") %>%
  bind_rows(
    final_water_sf_long_avg %>%
      select(year, water_stress) %>%
      rename(value = water_stress) %>%
      mutate(variable = "Water Stress")
  )

ggplot(final_combined_long_avg, aes(x = year, y = value, color = variable)) +
  geom_line() +
  geom_point() +
  labs(title = "Time Series of SPEI and Water Stress (2015-2020)",
       x = "Year",
       y = "Value") +
  theme_minimal()
```

## Task 15

Plot the time series of SPEI, water stress, and Agricultural GVA of each Serbian NUTS 3 region.

Both methodologies A and B can be used for this task

A. Plot the time series of SPEI, water stress, and Agricultural GVA of each Serbian NUTS 3 region.

Reshape in long form the dataset without avereging by regiones
```{r}
final_SPEI_sf_long <- final_sf %>%
  select(SPEI_2015, SPEI_2016, SPEI_2017, SPEI_2018, 
         SPEI_2019, SPEI_2020, NUTS_ID) %>%
  pivot_longer(cols = starts_with("SPEI_"), 
               names_to = "year", 
               values_to = "spei") 
```

Remove "SPEI_" from the year name and leave just the year (number)
```{r}
final_SPEI_sf_long <- final_SPEI_sf_long %>%
  mutate(year = as.numeric(gsub("SPEI_", "", year)))

final_water_sf_long <- final_sf %>%
  select(water_stress_2015, water_stress_2016, water_stress_2017,
         water_stress_2018, water_stress_2019, water_stress_2020, NUTS_ID) %>%
  pivot_longer(cols = starts_with("water_stress_"),
               names_to = "year",   
               values_to = "water_stress")
```

Remove " water_stress_" from the year name and leave just the year (number)
```{r}
final_water_sf_long <- final_water_sf_long %>%
  mutate(year = as.numeric(gsub("water_stress_", "", year)))

final_agr_sf_long <- final_sf %>%
  select(agr_gva_2015, agr_gva_2016, agr_gva_2017, agr_gva_2018, 
         agr_gva_2019, agr_gva_2020, NUTS_ID) %>%
  pivot_longer(cols = starts_with("agr_gva_"),
               names_to = "year",
               values_to = "agr_gva")
```

Remove "agr_gva_ from the year name and leave just the year (number)
```{r}
final_agr_sf_long <- final_agr_sf_long %>%
  mutate(year = as.numeric(gsub("agr_gva_", "", year)))
```

Merge the two datasets while preserving NUTS_ID
```{r}
final_combined_long <- final_SPEI_sf_long %>%
  select(NUTS_ID, year, spei) %>%  
  rename(value = spei) %>%
  mutate(variable = "SPEI") %>%
  bind_rows(
    final_water_sf_long %>%
      select(NUTS_ID, year, water_stress) %>% 
      rename(value = water_stress) %>%
      mutate(variable = "water stress")) %>%
  bind_rows(
    final_agr_sf_long %>%
      select(NUTS_ID, year, agr_gva) %>% 
      rename(value = agr_gva) %>%
      mutate(variable = "agr_gva")) %>%
  arrange(NUTS_ID)
```

generate the graph
```{r}
ggplot(final_combined_long, aes(x = year, y = value, color = variable)) +
  geom_line() +
  geom_point() +  
  facet_wrap(~ NUTS_ID, scales = "free_y") + 
  labs(title = "Grafici per NUTS_ID",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

COMMENT: 
SPEI and water stress overlaps
 
B. for a more clear representation: Plot the time series of each variable for every
NUTS3 separatelly and add the averege 

disaggregate final_combined_long for each variable
```{r}
final_combined_long_SPEI <- final_combined_long %>%
  filter( variable == "SPEI")
```

plot the SPEI time series and their mean 
```{r}
ggplot(final_combined_long_SPEI, aes(x = year, y = value, color = NUTS_ID)) +
  geom_line() +
  geom_line(data = final_SPEI_sf_long_avg, aes(x = year, y = spei),
            color = "black", linewidth = 2) +
  labs(title = "Time series for SPEI in each NUTS with Average SPEI (black)",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.title = element_blank())
```
COMMENT:
the time series in each NUTS of SPEI is quite similare over years.

disaggregate final_combined_long for each variable
```{r}
final_combined_long_water_stress <- final_combined_long %>%
  filter( variable == "water stress")
```


plot the water stress time series and their mean 
```{r}
ggplot(final_combined_long_water_stress,
       aes(x = year, y = value, color = NUTS_ID)) +
  geom_line() +
  geom_line(data = final_water_sf_long_avg, aes(x = year, y = water_stress),
            color = "black", linewidth = 2) +
  labs(title = "Time series for water stress in each NUTS with Average water stress (black)",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.title = element_blank())
```
COMMENT:
The water stress time series is quite heterogenous between serbian's NUTS3.

disaggregate final_combined_long for each variable
```{r}
final_combined_long_agr_gva <- final_combined_long %>%
  filter( variable == "agr_gva")
```

plot the agr_avg time series and their mean
```{r}
ggplot(final_combined_long_agr_gva, aes(x = year, y = value, color = NUTS_ID)) +
  geom_line() +
  labs(title = "Time series for agriculture in each NUTS",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.title = element_blank())
```

## Task 16

Plot a map showing the growth rate of SPEI from 2015 to 2020 of each Serbian
NUTS 3 region, in CRS 4326. 

compute the SPEI growth
```{r}
final_sf <- final_sf %>%
  mutate(speigr = (SPEI_2020 - SPEI_2015) / SPEI_2015 * 100) 
```

plot the growth of SPEI
```{r}
ggplot(data = final_sf) +
  geom_sf(aes(fill = speigr), color = "black") +
  scale_fill_viridis_c(option = "magma", name = "SPEI Growth Rate (%)") +  
  labs(title = "Growth Rate of SPEI from 2015 to 2020 in Serbian NUTS 3 Regions",
       subtitle = "Percentage Growth Rate") +
  theme_minimal() +
  theme(legend.position = "right")
```
JUST FOR THE SEEK OF REPRESENTATION:
Since the SPEI (Standardized Precipitation-Evapotranspiration Index) primarily indicates drought or moisture conditions, a color scale from "dry" colors (e.g.,pink or white) to "wet" colors (e.g., blues or purples) effectively conveys this concept.

## Task 17

Plot a map showing the growth rate of the water stress index from 2015 to 2020 
of each Serbian NUTS3 region, in CRS 32634.

Trasform the data and coumpute the water stress growth
```{r}
final_sf_32634 <- final_sf %>%
  st_transform(crs = 32634) %>%  # Imposta il CRS a 32634
  mutate(wsgr = (water_stress_2020 - water_stress_2015) / water_stress_2015 * 100) 
```

plot the gowth
```{r}
ggplot(data = final_sf_32634) +
  geom_sf(aes(fill = wsgr), color = "black") +
  scale_fill_viridis_c(option = "viridis", name = "water stress Growth Rate (%)") + 
  labs(title = "Growth Rate of water stress from 2015 to 2020 in Serbian NUTS 3 Regions",
       subtitle = "Percentage Growth Rate") +
  theme_minimal() +
  theme(legend.position = "right")
```
JUST FOR THE SEEK OF REPRESENTATION:
For a variable like water stress, which often represents a severity scale, a viridis or cividis color scale works well, as it allows for a gradient from low to high values without being overly intense.

COMMENT:
no great changing in scale turning the plot in CRS = 32634


## Task 18

Plot a map showing the growth rate of Agricultural GVA from 2015 to 2020 of each Serbian NUTS 3
region, in CRS 3035. 

Trasform the scale of the data and coumpute the agr_gva_ average
```{r}
final_sf_3035 <- final_sf %>%
  st_transform(crs = 3035) %>%  # Imposta il CRS a 3035
  mutate(agrgvagr = ( agr_gva_2020 - agr_gva_2015) / agr_gva_2015 * 100) 
```

plot
```{r}
ggplot(data = final_sf_3035) +
  geom_sf(aes(fill = agrgvagr), color = "black") +
  scale_fill_viridis_c(option = "viridis",
                       name = "Agricultural GVA Growth Rate (%)") +
  labs(title = "Growth Rate of Agricultural GVA from 2015 to 2020 in Serbian NUTS 3 Regions",
       subtitle = "Percentage Growth Rate") +
  theme_minimal() +
  theme(legend.position = "right")
```

plot the map again using blue for regions experiencing a negative growth, lightblue
for those having a growth rate between 0% and 5%, and white for regions with a growth rate >5%.

characterize each bin by a colour
```{r}
final_sf_3035 <- final_sf_3035 %>%
  mutate(agrvagr_category = case_when(
    agrgvagr < 0 ~ "Negative Growth",
    agrgvagr >= 0 & agrgvagr <= 5 ~ "0-5% Growth",
    agrgvagr > 5 ~ ">5% Growth"
  ))
```

generate a map
```{r}
ggplot(data = final_sf_3035) +
  geom_sf(aes(fill = agrvagr_category), color = "black") +  
  scale_fill_manual(values = c("Negative Growth" = "blue", 
                               "0-5% Growth" = "lightblue", 
                               ">5% Growth" = "white"),
                    name = "Agricultural GVA Growth Rate") +
  labs(title = "Growth Rate of Agricultural GVA from 2015 to 2020 in Serbian NUTS 3 Regions",
       subtitle = "Percentage Growth Rate") +
  theme_minimal() +
  theme(legend.position = "right")
```

##  Task 19  

Let’s define a new dataset with monthly frequency. The study area is Europe and 
the dataset shows consumption, investment and GDP per capita for each NUTS 3 
region from 2015 to 2020. Could GDP per capita in one region be affected by 
the lagged investment in a neighboring region? How could you check it? 

Yes, it could be affected by neighbors country investment and a way to check it is through 
spatial autocorrelation tests such as Moran I test and generally speaking 
apply a spatial econometric approach that accounts for potential spatial dependencies over time.

The several steps involve:
- Dataset Setup: Ensure monthly data from 2015 to 2020 for GDP per capita, consumption, and investment in each NUTS 3 region
- Lagged Variables: Create a lagged spatial variable for investment, which incorporates investment values from neighboring regions with a time lag.
- Modeling Approach:
    SAR: Uses a spatial lag of GDP to capture dependency on neighboring regions.
    SDM: Tests both local effects and spillover effects from neighboring investments.
       Testing Spatial Dependence: Use Moran’s I and Lagrange Multiplier tests to detect spatial           autocorrelation, and consider spatial panel models for temporal and spatial effects.


# SECTION B : SECOND NOT MANDATORY TASKS

Using the “queen” criterion = two areas are considered affecting each other if they share a border

compute the Moran’s I Test for SPEI in 2015, and for Agricultural GVA in 2015
```{r}
final_sf <- st_make_valid(final_sf)
final_sf <- st_transform(final_sf, crs = 4326)
```

Spatial Lags
```{r}
k <- poly2nb(final_sf, row.names = final_sf$NUTS_ID)
k_weights <- nb2listw(k, style = "W") 
```

create centroids
```{r}
coords <- st_coordinates(st_centroid(st_geometry(final_sf)))
p <- plot(k, coords)
```

create Spatial Lags variable
```{r}
index_lag <- lag.listw(k_weights, final_sf$agr_gva_2015)
```

Is population spatially correlated across Europe?

Perform the Moran's I Test and plot of autocorrelation 

Moran's I Test
```{r}
moran1_agr2015 <- moran.test(final_sf$agr_gva_2015,k_weights)
print(moran1_agr2015) 
```
COMMENT: 
The Moran's I test result of 0.58 indicates a moderately strong positive spatial autocorrelation.
The spatial autocorrelation is statistically significant given a p-value of 4.234e-07.

create a spatial autocorrelation scatterplot for the above-mentioned variables

Precompute scaled values
```{r}
final_sf <- final_sf %>%
  mutate(agr_gva_2015_scaled = scale(agr_gva_2015),
         index_lag_scaled = scale(index_lag))
```

Compute mean values for horizontal and vertical lines
```{r}
mean_agr_gva_2015_scaled <- mean(final_sf$agr_gva_2015_scaled, na.rm = TRUE)
mean_index_lag_scaled <- mean(final_sf$index_lag_scaled, na.rm = TRUE)
```

Create the plot
```{r}
ggplot <- ggplot(final_sf) +
  geom_point(aes(x = agr_gva_2015_scaled, y = index_lag_scaled, fill = NUTS_ID),
             shape = 22, size = 3, alpha = 0.8) +
  geom_hline(yintercept = mean_index_lag_scaled, linetype = "dashed") +
  geom_vline(xintercept = mean_agr_gva_2015_scaled, linetype = "dashed") +
  geom_smooth(aes(x = agr_gva_2015_scaled, y = index_lag_scaled),
              method = "lm", color = "black") +
  labs(x = "agr_gva_2015 (scaled)", y = "Lagged Index (scaled)",
       fill = "Country") +
  theme_classic()
ggplot
```
Can the Moran’s I Test be used to test the effect of a lagged variable on the dependent variable in a neighboring region? For instance, the effect of investment at t-1 in region j on gdp at t in region i? 

While Moran’s I can check for spatial autocorrelation, a spatial regression model is necessary to assess specific spatial lagged effects.
The Moran’s I test measures spatial autocorrelation, indicating whether a variable  (e.g., GDP per capita) is clustered, dispersed, or randomly distributed across regions.
However, it doesn’t directly test the causal impact of a lagged variable in neighboring regions 
(like previous investment in one region on GDP in another).
To explore the effect of lagged investment in neighboring regions on GDP, a spatial lag model (SLM) 
or a spatial Durbin model (SDM) would be more appropriate. These models incorporatespatially lagged independent variables, allowing for analysis of the potential impact of investment at t−1 in region j on GDP at t in region i.


# SECTION B : THIR NOT MANDATORY TASK

Download Aqueduct gdb data and save the dataset as a shp file

First check layers and choose just the layer "baseline_monthly"
```{r}
layers <- st_layers("data/Aqueduct40_waterrisk/GDB/Aq40_Y2023D07M05.gdb")
print(layers)
```

load data for the baseline_annual layer
```{r}
gdb_path <-"data/Aqueduct40_waterrisk/GDB/Aq40_Y2023D07M05.gdb"
Aqueduct_gdb <- st_read(gdb_path, layer = "baseline_annual")
```
Filter the data for Serbia using 'gid_0' column (assuming Serbia is labeled as "SRB" in 'gid_0')
```{r}
serbia_aqua_sf <- Aqueduct_gdb %>%
  filter(gid_0 == "SRB")
```

Remove any rows with missing data to reduce computational burden
```{r}
serbia_aqua_sf <- serbia_aqua_sf %>%
  drop_na()
```

Crop the Aqueduct data to Serbia's NUTS 3 shapefile shp3
```{r}
serbia_aqua_cropped <- st_intersection(serbia_aqua_sf, shp3)
#Plot data
plot(st_geometry(serbia_aqua_cropped))
```


Save the dataset filtered as a shape file
```{r}
# Specify the output path in the 'data' directory
output_path <- "data/Aqueduct_Serbia.shp"

# Delete the files if they already exist
if (file.exists(output_path)) {
  file.remove(output_path)
  file.remove(sub(".shp$", ".shx", output_path))  # Remove associated .shx file
  file.remove(sub(".shp$", ".dbf", output_path))  # Remove associated .dbf file
  file.remove(sub(".shp$", ".prj", output_path))  # Remove associated .prj file (if exists)
}

# Save the file
# st_write(serbia_aqua_cropped, "serbia_aqueduct_data.shp", delete_layer = TRUE)
```

